---
title: "Teacher Descriptive Stats"
author: "Evan Yacek"
date: "Tuesday, July 12, 2016"
output: html_document
---

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.align="center", fig.height=6, fig.width=6)
library(ggplot2)
theme_set(theme_bw(base_size=12))
library(tidyr)
library(plyr)
library(dplyr)
library(nhanesA)
```


```{r, echo=FALSE , include=FALSE}


OCP_E2007 <- nhanes('OCQ_E')
OCP_F2009 <- nhanes('OCQ_F')
DEMO2007 <- nhanes('DEMO_E')
DEMO2009 <- nhanes('DEMO_F')
Diet2007 <- nhanes('DBQ_E')
Diet2009 <- nhanes('DBQ_F')
cons2007 <- nhanes('CBQ_E')
cons2009 <- nhanes('CBQ_F')
bodyMeasure2007 <- nhanes('BMX_E')
bodyMeasure2009 <- nhanes('BMX_F')
bp2007 <- nhanes('BPX_E')
bp2009 <- nhanes('BPX_F')
chol2007 <- nhanes('HDL_E')
chol2009 <- nhanes('HDL_F')
tri2007 <- nhanes('TRIGLY_E')
tri2009 <- nhanes('TRIGLY_F')
apob2007 <-nhanes('APOB_E')
apob2009 <-nhanes('APOB_F')

#FILTER BASED ON TEACHER OCCUPATION CODES
TEACHERS2007 <- filter(OCP_E2007, OCD231 == 15 & OCD241 == 8 )
TEACHERS2009 <- filter(OCP_F2009, OCD231 == 15 & OCD241 == 8 )

#MERGE INTO SINGLE TEACHER DATASET
TEACHERDEMO2007 <- merge(TEACHERS2007 , DEMO2007) 
td <- merge(TEACHERDEMO2007 , Diet2007)
td <- merge(td , cons2007)
td <- merge(td, bp2007)
td <- merge(td, bodyMeasure2007)
td <- merge(td, chol2007)



TEACHERDEMO2009 <- merge(TEACHERS2009 , DEMO2009)
td2 <- merge(TEACHERDEMO2009 , Diet2009)
td2 <- merge(td2, cons2009)
td2 <- merge(td2, bp2009)
td2 <- merge(td2, bodyMeasure2009)
td2 <- merge(td2, chol2009)



common_cols <- intersect(colnames(td), colnames(td2))
newCombine<- 
  rbind(
  subset(td, select = common_cols), 
  subset(td2, select = common_cols)
)



#CREATE FREQUENCY TABLE FOR GENDER
MF_Frequency <- table(newCombine$RIAGENDR)  
res <- cbind(MF_Frequency,round(prop.table(MF_Frequency)*100, 2))
rownames(res) <- c('Male','Female')
colnames(res) <- c("Freq", "Pct")

```
### Teacher's Gender Distribution Frequency Table(N= 279)
```{r, echo = FALSE}
res
```
### Intervaled Age Distribution For Teachers(N = 279)
```{r, echo= FALSE}

#INTERVALED AGE DISTRIBUTION FOR TEACHERS

x <- newCombine$RIDAGEYR
factorx <- factor(cut(x, breaks=8))
#Tabulate and turn into data.frame
xout <- as.data.frame(table(factorx))
#Add cumFreq and proportions
xout <- transform(xout, cumFreq = cumsum(Freq), relative = prop.table(Freq)*100)
colnames(xout) <- c("Age-Range" , "Freq" , "Total", "PCT")
xout
ggplot(newCombine, aes(x = RIDAGEYR, fill=..count.. )) + geom_histogram(bins = 8, col="black" ) + labs(title="Distribution for Age of Teachers") + labs(x="Age", y="Count") + scale_fill_gradient("Count", low = "green", high = "red")

ageDataset2007 <- merge(DEMO2007, cons2007)
ageDataset2007 <- merge(ageDataset2007, Diet2007 )
ageDataset2009 <- merge(DEMO2009, cons2009)
ageDataset2009 <- merge(ageDataset2009, Diet2009)
common_cols5 <- intersect(colnames(ageDataset2007), colnames(ageDataset2009))
ageDataset<- 
  rbind(
  subset(ageDataset2007, select = common_cols5), 
  subset(ageDataset2009, select = common_cols5)
)
ageDataset <- filter(ageDataset, RIDAGEYR >= 16 & RIDAGEYR <= 80 )
```
### Intervaled (16 < Age < 90) Distribution for NHANES(07-10) N = 13435
```{r, echo=FALSE}
y <- ageDataset$RIDAGEYR
factory <- factor(cut(y, breaks=8))
#Tabulate and turn into data.frame
yout <- as.data.frame(table(factory))
#Add cumFreq and proportions
yout <- transform(yout, cumFreq = cumsum(Freq), relative = prop.table(Freq)*100)
colnames(yout) <- c("Age-Range" , "Freq" , "Total", "PCT")
yout
ggplot(ageDataset, aes(x = RIDAGEYR, fill=..count.. )) + geom_histogram(bins = 8, col="black" ) + labs(title="Distribution for Age of Population NHANES(07-10) N = 13435") + labs(x="Age", y="Count") + scale_fill_gradient("Count", low = "green", high = "red")
```
### Teacher Ethnicity Distribution

```{r, echo=FALSE}
#Ethnicity Distribution
Ethnic_Frequency <- table(newCombine$RIDRETH1)  
rest <- cbind(Ethnic_Frequency,round(prop.table(Ethnic_Frequency)*100, 2))
colnames(rest) <- c("Freq", "Pct")
rownames(rest) <- c("Mexican American", "Other Hispanic", "Non-Hispanic White", "Non-Hispanic Black", "Other")
rest

```


### Male Teachers mean waist circumference N= 79

```{r , echo = FALSE}

#Calculate mean Waist Cirucumference 
#Seperate Men and Women
maleTeachers <- filter(newCombine,RIAGENDR == 1 )
femTeachers <- filter(newCombine,RIAGENDR == 2 )
maleTeacherWaistMean <- mean(maleTeachers$BMXWAIST, na.rm = TRUE)

#Average
maleTeacherWaistMean
```
### Female mean waist circumference N = 200
```{r, echo= FALSE}
femTeacherWaistMean <- mean(femTeachers$BMXWAIST, na.rm = TRUE)
femTeacherWaistMean
totalPop <- rbind(DEMO2007, DEMO2009)
totalPop2 <- rbind(bodyMeasure2007, bodyMeasure2009)
totalPopset <- merge(totalPop, totalPop2)
totalPopset <- filter(totalPopset, RIDAGEYR >= 16 & RIDAGEYR <= 80 )
totalPopMale <- filter(totalPopset, RIAGENDR == 1)
totalPopFemale <- filter(totalPopset, RIAGENDR == 2)
femPopMeanWaist <- mean(totalPopFemale$BMXWAIST, na.rm = TRUE)
malePopMeanWaist <- mean(totalPopMale$BMXWAIST, na.rm = TRUE)
```
### Female NHANES(07-10) mean waist circumference N = 6618
```{r, echo=FALSE}
femPopMeanWaist
```

### Male NHANES(07-10) mean waist circumference N = 6406
```{r,echo = FALSE}
malePopMeanWaist



```

### Mean triglyceride levels for Male Teachers N= 42


```{r , echo=FALSE}

teacherTri2007 <- filter(OCP_E2007, OCD231 == 15 & OCD241 == 8 )
teacherTri2007 <- merge(teacherTri2007, tri2007 )
teacherTri2007 <- merge(teacherTri2007, DEMO2007 )
teacherTri2009 <- filter(OCP_F2009, OCD231 == 15 & OCD241 == 8 )
teacherTri2009 <- merge(teacherTri2009, tri2009 )
teacherTri2009 <- merge(teacherTri2009, DEMO2009 )

common_cols2 <- intersect(colnames(teacherTri2007), colnames(teacherTri2009))
newCombine2<- 
  rbind(
  subset(teacherTri2007, select = common_cols2), 
  subset(teacherTri2009, select = common_cols2)
)


maleTeachTri <- filter(newCombine2 ,RIAGENDR == 1 )
maleTeachTriMean <- mean(maleTeachTri$LBXTR, na.rm = TRUE)
maleTeachTriMean
```

### Mean triglyceride levels for Female Teachers N= 87

```{r, echo = FALSE}
femaleTeachTri <- filter(newCombine2 ,RIAGENDR == 2 )
femaleTeachTriMean <- mean(femaleTeachTri$LBXTR, na.rm = TRUE)
femaleTeachTriMean
```

```{r, echo=FALSE}

totalPop3 <- rbind(DEMO2007, DEMO2009)
totalPop4 <- rbind(tri2007, tri2009)
totalPopset2 <- merge(totalPop3, totalPop4)
totalPopset2 <- filter(totalPopset2, RIDAGEYR >= 16 & RIDAGEYR <= 80 )
totalPopMale2 <- filter(totalPopset2, RIAGENDR == 1)
totalPopFemale2 <- filter(totalPopset2, RIAGENDR == 2)
```
### Mean Triglyceride for NHANES Female pop. N = 3502
```{r, echo=FALSE}
femPopMeanTri <- mean(totalPopFemale2$LBXTR, na.rm = TRUE)
femPopMeanTri
```
### Mean Triglyceride for NHANES Male pop. N = 3394
```{r, echo = FALSE}
malePopMeanTri <- mean(totalPopMale2$LBXTR, na.rm = TRUE)
malePopMeanTri

```

### Mean HDL for Male Teachers N = 79

```{r, echo = FALSE}

teacherHDL2007 <- filter(OCP_E2007, OCD231 == 15 & OCD241 == 8 )
teacherHDL2007 <- merge(teacherHDL2007, chol2007 )
teacherHDL2007 <- merge(teacherHDL2007, DEMO2007 )
teacherHDL2009 <- filter(OCP_F2009, OCD231 == 15 & OCD241 == 8 )
teacherHDL2009 <- merge(teacherHDL2009, chol2009 )
teacherHDL2009 <- merge(teacherHDL2009, DEMO2009 )

common_cols3 <- intersect(colnames(teacherHDL2007), colnames(teacherHDL2009))
newCombine3<- 
  rbind(
  subset(teacherHDL2007, select = common_cols3), 
  subset(teacherHDL2009, select = common_cols3)
)


maleTeachHDL <- filter(newCombine3 ,RIAGENDR == 1 )
maleTeachHDLMean <- mean(maleTeachHDL$LBDHDD, na.rm = TRUE)
maleTeachHDLMean
```

### Mean HDL for Female Teachers N = 200
```{r, echo= FALSE}

femaleTeachHDL <- filter(newCombine3 ,RIAGENDR == 2 )
femaleTeachHDLMean <- mean(femaleTeachHDL$LBDHDD, na.rm = TRUE)
femaleTeachHDLMean


```

### Mean HDL for Female NHANES Pop. N = 6618
```{r, echo = FALSE}
ageHDL <- rbind(chol2007, chol2009)
ageHDL2 <- rbind(DEMO2007, DEMO2009)
ageHDL3 <- merge(ageHDL, ageHDL2)
ageHDL3 <- filter(ageHDL3, RIDAGEYR >= 16 & RIDAGEYR <= 80 )
femAgeHDL <- filter( ageHDL3, RIAGENDR == 2 )
maleAgeHDL <- filter( ageHDL3, RIAGENDR == 1 )
femAgeMean <- mean(femAgeHDL$LBDHDD, na.rm = TRUE)
femAgeMean
```

### Mean HDL for Male NHANES Pop. N = 6406

```{r, echo=FALSE}
maleAgeMean <- mean( maleAgeHDL$LBDHDD , na.rm = TRUE)
maleAgeMean
```

```{r , echo= FALSE}

TEACHERS2007 <- merge(TEACHERS2007, Diet2007)
TEACHERS2009 <- merge(TEACHERS2009, Diet2009)
commoncols <- intersect(colnames(TEACHERS2007), colnames(TEACHERS2009))
teacherCombine<- 
  rbind(
    subset(TEACHERS2007, select = commoncols), 
    subset(TEACHERS2009, select = commoncols)
  )

healthyDietQ <- as.data.frame(table(teacherCombine$DBQ700))
healthyDietQ$Var1 <- c("Excellent","Very good", "Good", "Fair", "Poor")
positions <- c("Excellent","Very good", "Good", "Fair", "Poor")
ggplot(healthyDietQ, aes(x=Var1, y = Freq*100/288)) + geom_bar(stat="identity" , fill = "sky blue")+
  geom_text(aes(label = sprintf("%.2f%%", Freq/sum(Freq) * 100)), vjust = -.5)+ labs(title = "How healthy is your diet?(Total) N= 288", x = "Rating(1-5)", y = "PCT") + scale_x_discrete(limits = positions)
  
healthyDietQ2007 <- as.data.frame(table(TEACHERS2007$DBQ700))
healthyDietQ2007$Var1 <- c("Excellent","Very good", "Good", "Fair", "Poor")
positions <- c("Excellent","Very good", "Good", "Fair", "Poor")
ggplot(healthyDietQ2007, aes(x=Var1, y = Freq*100/144)) + geom_bar(stat="identity" , fill = "sky blue")+
  geom_text(aes(label = sprintf("%.2f%%", Freq/sum(Freq) * 100)), vjust = -.5)+ labs(title = "How healthy is your diet?(2007-08) N= 144", x = "Rating(1-5)", y = "PCT")+ scale_x_discrete(limits = positions)

healthyDietQ2009 <- as.data.frame(table(TEACHERS2009$DBQ700))
healthyDietQ2009$Var1 <- c("Excellent","Very good", "Good", "Fair" )
positions <- c("Excellent","Very good", "Good", "Fair" )
ggplot(healthyDietQ2009, aes(x=Var1, y = Freq*100/144)) + geom_bar(stat="identity" , fill = "sky blue")+
  geom_text(aes(label = sprintf("%.2f%%", Freq/sum(Freq) * 100)), vjust = -.5)+ labs(title = "How healthy is your diet?(2009-10) N= 144", x = "Rating(1-5)", y = "PCT") +  scale_x_discrete(limits = positions)


healthyDietQPOP<- as.data.frame(table(ageDataset$DBQ700))
healthyDietQPOP$Var1 <- c("Excellent","Very good", "Good", "Fair", "Poor" ,"IDK", "NA")
positions <- c("Excellent","Very good", "Good", "Fair", "Poor" ,"IDK", "NA")
ggplot(healthyDietQPOP, aes(x=Var1, y = Freq*100/144)) + geom_bar(stat="identity" , fill = "sky blue")+
  geom_text(aes(label = sprintf("%.2f%%", Freq/sum(Freq) * 100)), vjust = -.5)+ labs(title = "How healthy is your diet? N= 13,435", x = "Rating(1-7)", y = "PCT") + scale_x_discrete(limits = positions)

TEACHERS2007 <- merge(TEACHERS2007, cons2007)
TEACHERS2009 <- merge(TEACHERS2009, cons2009)
commoncols <- intersect(colnames(TEACHERS2007), colnames(TEACHERS2009))
teacherCombine<- 
  rbind(
    subset(TEACHERS2007, select = commoncols), 
    subset(TEACHERS2009, select = commoncols)
  )

fruitsQ <- as.data.frame(table(teacherCombine$CBQ020))
fruitsQ$Var1 <- c("Always","Most of the \n Time", "Sometimes", "Rarely")
positions <- c("Always","Most of the \n Time", "Sometimes", "Rarely")
ggplot(fruitsQ, aes(x=Var1, y = Freq*100/287)) + geom_bar(stat="identity" , fill = "red")+
  geom_text(aes(label = sprintf("%.2f%%", Freq/sum(Freq) * 100)), vjust = -.5)+ labs(title = "How often do you have fruits \n in your home?(Total) N= 287", x = "Rating(1-7)", y = "PCT")+ scale_x_discrete(limits = positions)

fruitsQPOP <- as.data.frame(table(ageDataset$CBQ020))
fruitsQPOP$Var1 <- c("Always","Most of the \n Time", "Sometimes", "Rarely" ,"Never", "IDK" , "NA")
positions <- c("Always","Most of the \n Time", "Sometimes", "Rarely" ,"Never", "IDK" , "NA")
ggplot(fruitsQPOP, aes(x=Var1, y = Freq*100/287)) + geom_bar(stat="identity" , fill = "red")+
  geom_text(aes(label = sprintf("%.2f%%", Freq/sum(Freq) * 100)), vjust = -.5)+ labs(title = "How often do you have fruits \n in your home?(Total) N= 13,435", x = "Rating(1-7)", y = "PCT")  + scale_x_discrete(limits = positions)

vegQ <- as.data.frame(table(teacherCombine$CBQ030))
vegQ$Var1 <- c("Always","Most of the \n Time", "Sometimes", "Rarely" , "Never")
positions <- c("Always","Most of the \n Time", "Sometimes", "Rarely" , "Never")
ggplot(vegQ, aes(x=Var1, y = Freq*100/287)) + geom_bar(stat="identity" , fill = "dark green")+
  geom_text(aes(label = sprintf("%.2f%%", Freq/sum(Freq) * 100)), vjust = -.5)+ labs(title = "How often do you have dark green vegetables \n in your home?(Total) N= 287", x = "Rating(1-7)", y = "PCT")+ scale_x_discrete(limits = positions)

vegQPOP <- as.data.frame(table(ageDataset$CBQ030))
vegQPOP$Var1 <- c("Always","Most of the \n Time", "Sometimes", "Rarely" ,"Never", "IDK" , "NA")
positions <- c("Always","Most of the \n Time", "Sometimes", "Rarely" ,"Never", "IDK" , "NA")
ggplot(vegQPOP, aes(x=Var1, y = Freq*100/287)) + geom_bar(stat="identity" , fill = "dark green")+
  geom_text(aes(label = sprintf("%.2f%%", Freq/sum(Freq) * 100)), vjust = -.5)+ labs(title = "How often do you have dark green vegetables \n in your home?(Total) N= 13,435", x = "Rating(1-7)", y = "PCT")  + scale_x_discrete(limits = positions)

saltQ <- as.data.frame(table(teacherCombine$CBQ040))
saltQ$Var1 <- c("Always","Most of the \n Time", "Sometimes", "Rarely" , "Never", "IDK")
positions <- c("Always","Most of the \n Time", "Sometimes", "Rarely" , "Never", "IDK")
ggplot(saltQ, aes(x=Var1, y = Freq*100/287)) + geom_bar(stat="identity" , fill = "blue")+
  geom_text(aes(label = sprintf("%.2f%%", Freq/sum(Freq) * 100)), vjust = -.5)+ labs(title = "How often do you have salty snacks(chips/crackers) \n in your home?(Total) N= 287", x = "Rating(1-7)", y = "PCT")+ scale_x_discrete(limits = positions)

saltQPOP <- as.data.frame(table(ageDataset$CBQ040))
saltQPOP$Var1 <- c("Always","Most of the \n Time", "Sometimes", "Rarely" ,"Never", "IDK" , "NA")
positions <- c("Always","Most of the \n Time", "Sometimes", "Rarely" ,"Never", "IDK" , "NA")
ggplot(saltQPOP, aes(x=Var1, y = Freq*100/287)) + geom_bar(stat="identity" , fill = "blue")+
  geom_text(aes(label = sprintf("%.2f%%", Freq/sum(Freq) * 100)), vjust = -.5)+ labs(title = "How often do you have Salty Snacks \n in your home?(Total) N= 13,435", x = "Rating(1-7)", y = "PCT")  + scale_x_discrete(limits = positions)



softQ <- as.data.frame(table(teacherCombine$CBQ060))
softQ$Var1 <- c("Always","Most of the \n Time", "Sometimes", "Rarely" , "Never", "IDK")
positions <- c("Always","Most of the \n Time", "Sometimes", "Rarely" , "Never", "IDK")
ggplot(softQ, aes(x=Var1, y = Freq*100/287)) + geom_bar(stat="identity" , fill = "purple")+
  geom_text(aes(label = sprintf("%.2f%%", Freq/sum(Freq) * 100)), vjust = -.5)+ labs(title = "How often do you have soft drinks \n available in your home?(Total) N= 287", x = "Rating(1-7)", y = "PCT")+ scale_x_discrete(limits = positions)


softQPOP <- as.data.frame(table(ageDataset$CBQ060))
softQPOP$Var1 <- c("Always","Most of the \n Time", "Sometimes", "Rarely" ,"Never", "IDK" , "NA")
positions <- c("Always","Most of the \n Time", "Sometimes", "Rarely" ,"Never", "IDK" , "NA")
ggplot(softQPOP, aes(x=Var1, y = Freq*100/287)) + geom_bar(stat="identity" , fill = "purple")+
  geom_text(aes(label = sprintf("%.2f%%", Freq/sum(Freq) * 100)), vjust = -.5)+ labs(title = "How often do you have Soft Drinks \n in your home?(Total) N= 13,435", x = "Rating(1-7)", y = "PCT")  + scale_x_discrete(limits = positions)
```




